---
title: "Initial Korean covid 19"
author: "Alfredo Acevedo"
date: "3/19/2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?

## Methods
From what we got from the problem, it can be solved as categorical - categorical. so taking these two to solve for the answer. Which is age and sex as the two category.




## Graphical results - use ggplot commands to do this  
  

```{r}
dd2 <- kRdata %>% group_by(result,sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
According to the graph it shows that women appear to be getting covid more then men. but men appear to die to covid more.  
  





  
## Numerical results
  
```{r}
table2 <- xtabs(~result + sex, data=kRdata)
rowPerc(table2)
colPerc(table2)
```
  
It still shows that women get covid more but die less. For the men they get covid less then women but are more likely to die from covid. 

## Inferential results 
  
```{r}
chisq.test(table2)
chisqtestGC(table2)
```
  
```{r}
fisherXY(kRdata$sex,kRdata$result)
```
  




## Conclusions
With the data above from determining the death rates and the rate of getting covid. we can confirm that the more females got covid, but are least likely to die. but males are more likely to die covid, but are 
